### 2025-03-15 docs/manuscript/manuscript.tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[numbers,sort&compress]{natbib}
\usepackage[a4paper,left=3cm,right=3cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{setspace}
\usepackage{times}
\usepackage{authblk}
\usepackage{lineno}
\usepackage{caption}

% Formatting settings
\onehalfspacing
\captionsetup{font=small,labelfont=bf}
\renewcommand\Affilfont{\itshape\small}

% Title formatting
\makeatletter
\renewcommand\maketitle{
  \begin{flushleft}
  {\LARGE\bfseries\@title\par}
  \vskip 1em
  {\large\@author\par}
  \vskip 0.5em
  {\small\@date}
  \end{flushleft}
  \par\vskip 1em
}
\makeatother

\title{Package creation for understanding how the frequency of lab tests obtained vary dependent on objective clinical conditions.}

\author[1]{Author One}
\author[2]{Author Two}
\affil[1]{Affiliation One}
\affil[2]{Affiliation Two}

\date{\today}

\begin{document}
\linenumbers
\maketitle

\begin{abstract}
% Abstract content goes here
\end{abstract}

\section{Introduction}
% Introduction content goes here

\section{Methods}
% Methods content goes here

\section{Results}
% Results content goes here

\section{Discussion}
% Discussion content goes here

\section{Conclusion}
% Conclusion content goes here

\bibliographystyle{unsrtnat}
\bibliography{references}

\end{document}