### 2025-03-29 docs/manuscript/manuscript.tex
V4.

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[numbers,sort&compress]{natbib}
\usepackage[a4paper,left=3cm,right=3cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{setspace}
\usepackage{times}
\usepackage{authblk}
\usepackage{lineno}
\usepackage{caption}

% Formatting settings
\onehalfspacing
\captionsetup{font=small,labelfont=bf}
\renewcommand\Affilfont{\itshape\small}

% Title formatting
\makeatletter
\renewcommand\maketitle{
  \begin{flushleft}
  {\LARGE\bfseries\@title\par}
  \vskip 1em
  {\large\@author\par}
  \vskip 0.5em
  {\small\@date}
  \end{flushleft}
  \par\vskip 1em
}
\makeatother

\title{Care Phenotypes: A Novel Approach to Understanding Healthcare Data Collection Patterns}

\author[1]{Author One}
\author[2]{Author Two}
\affil[1]{Affiliation One}
\affil[2]{Affiliation Two}

\date{\today}

\begin{document}
\linenumbers
\maketitle

\begin{abstract}
Healthcare data collection patterns, particularly in laboratory measurements, often exhibit significant variation across patients that cannot be fully explained by objective clinical factors. This variation, which may reflect subjective decisions by medical staff, can introduce systematic biases in healthcare datasets and affect the validity of research findings. We present a novel approach to understanding these variations through the concept of "care phenotypes" - objective labels based on observable care patterns that reflect how patients are monitored and treated. We develop a Python package that enables researchers to identify and analyze these care phenotypes, accounting for legitimate clinical factors while highlighting unexplained variations in care delivery. Using examples from the MIMIC dataset, we demonstrate how care phenotypes can help researchers understand potential biases in their data and develop more robust healthcare algorithms. Our approach moves beyond traditional demographic labels for fairness evaluation, focusing instead on observable care patterns that may better reflect disparities in healthcare delivery.
\end{abstract}

\section{Introduction}

Healthcare datasets, particularly those derived from electronic health records (EHRs), have become invaluable resources for medical research and the development of healthcare algorithms. However, these datasets often contain systematic variations in data collection patterns that can significantly impact research validity and algorithmic fairness. This variation is particularly evident in laboratory measurements and routine care procedures, where the frequency and consistency of data collection can vary substantially across patients.

\subsection{The Challenge of Data Collection Variation}

In intensive care settings, for example, patients with similar objective measures of illness severity (such as SOFA scores or Charlson comorbidity indices) may receive markedly different frequencies of monitoring and testing. While some of this variation can be explained by legitimate clinical factors - such as illness severity or pre-existing conditions - significant unexplained variations often remain. These variations may reflect subjective decisions by medical staff about monitoring intensity, potentially introducing systematic biases into healthcare datasets.

\subsection{Current Limitations in Fairness Evaluation}

Traditional approaches to evaluating healthcare algorithm fairness often rely on demographic labels (race, ethnicity, gender) that may be poorly captured in healthcare data and may not fully reflect the complex factors influencing care decisions. These demographic-based approaches can miss important disparities in care delivery that manifest through variations in monitoring and treatment patterns.

\subsection{Introducing Care Phenotypes}

We propose a novel approach to understanding healthcare disparities through the concept of "care phenotypes" - objective labels based on observable care patterns that reflect how patients are monitored and treated. These phenotypes are derived from easily measurable metrics such as:
\begin{itemize}
    \item Frequency of laboratory measurements
    \item Regularity of routine care procedures
    \item Consistency of vital sign monitoring
\end{itemize}

\subsection{Objectives}

The primary objectives of this work are to:
\begin{itemize}
    \item Develop a framework for identifying and analyzing care phenotypes in healthcare datasets
    \item Create tools to help researchers understand potential biases in their data
    \item Provide methods for accounting for legitimate clinical factors while highlighting unexplained variations
    \item Enable more objective fairness evaluation of healthcare algorithms
\end{itemize}

\subsection{Implementation}

We present a Python package that implements this framework, focusing on:
\begin{itemize}
    \item Analysis of measurement frequencies and patterns
    \item Adjustment for clinical factors
    \item Creation of care phenotype labels
    \item Evaluation of healthcare algorithm fairness using these phenotypes
\end{itemize}

\section{Methods}

\subsection{Data Processing Framework}

We developed a comprehensive framework for processing MIMIC-IV data, implemented as a Python package. The framework consists of several key components:

\subsubsection{Data Structures and Formats}

We defined standardized data structures for various MIMIC data types, including:
\begin{itemize}
    \item Patient demographics and admission information
    \item Laboratory measurements and chart events
    \item ICU stays and clinical scores
\end{itemize}

These structures ensure type safety and consistency throughout the data processing pipeline. We implemented robust data validation and integrity checks to maintain data quality.

\subsubsection{Clinical Score Calculations}

Our framework includes implementations of several widely-used clinical scoring systems:

\begin{itemize}
    \item \textbf{SOFA (Sequential Organ Failure Assessment)}: Evaluates organ dysfunction across six systems (respiratory, coagulation, liver, cardiovascular, CNS, and renal)
    \item \textbf{Charlson Comorbidity Index}: Assesses patient comorbidity burden using 17 weighted conditions
    \item \textbf{APACHE II}: Comprehensive scoring system incorporating acute physiology, chronic health, and age components
    \item \textbf{SAPS II}: Simplified acute physiology scoring system
    \item \textbf{Elixhauser Comorbidity Index}: Detailed assessment of 31 comorbidities
\end{itemize}

Each scoring system is implemented as a modular component, allowing for flexible integration and extension. The implementations handle missing data gracefully and provide detailed component-level analysis.

\subsubsection{Data Processing Pipeline}

The data processing pipeline includes:
\begin{itemize}
    \item Standardized data loading and validation
    \item Automated data cleaning and normalization
    \item Efficient handling of time-series data
    \item Integration of multiple data sources
    \item Comprehensive error handling and logging
\end{itemize}

\subsection{Testing and Validation Framework}

We implemented a comprehensive testing and validation framework to ensure the reliability and robustness of our implementation. This framework consists of several key components:

\subsubsection{Synthetic Data Generation}

We developed a sophisticated synthetic data generator that creates MIMIC-like datasets for testing purposes:
\begin{itemize}
    \item Generation of realistic patient demographics and admission information
    \item Creation of synthetic laboratory measurements and chart events
    \item Simulation of ICU stays and clinical scores
    \item Preservation of temporal relationships and data dependencies
\end{itemize}

\subsubsection{Component-Level Testing}

Our testing framework includes comprehensive tests for each major component:
\begin{itemize}
    \item \textbf{Clinical Score Validation}: Verification of SOFA, Charlson, and other clinical score calculations
    \item \textbf{Data Processing Validation}: Testing of data cleaning, transformation, and integration
    \item \textbf{Result Validation}: Verification of phenotype creation and analysis results
\end{itemize}

\subsubsection{Integration Testing}

We implemented integration tests to validate the interaction between components:
\begin{itemize}
    \item End-to-end testing of the complete data processing pipeline
    \item Validation of data relationships and consistency
    \item Testing of error handling and edge cases
\end{itemize}

\subsubsection{Performance Testing}

Our framework includes comprehensive performance testing:
\begin{itemize}
    \item \textbf{Large Dataset Handling}: Testing with datasets of varying sizes (up to 10,000 patients)
    \item \textbf{Memory Usage Optimization}: Monitoring and optimization of memory consumption
    \item \textbf{Processing Speed Optimization}: Evaluation of processing time and scalability
\end{itemize}

\subsection{Core Functionality Implementation}

\subsubsection{Clinical Factor Adjustment}

We implemented a robust system for adjusting care patterns based on clinical factors:
\begin{itemize}
    \item Regression-based adjustment for multiple clinical factors
    \item Handling of missing values and outliers
    \item Preservation of data structure and relationships
    \item Comprehensive logging and error tracking
\end{itemize}

\subsubsection{Pattern Analysis}

Our implementation includes sophisticated methods for analyzing care patterns:
\begin{itemize}
    \item \textbf{Pattern Consistency}: Evaluation of care pattern stability across different measures
    \item \textbf{Unexplained Variation}: Quantification of variations not explained by clinical factors
    \item \textbf{Data Quality}: Comprehensive validation of pattern integrity
\end{itemize}

\subsubsection{Data Preprocessing}

We implemented a comprehensive suite of data preprocessing methods:
\begin{itemize}
    \item \textbf{Missing Value Handling}: Multiple strategies including mean, median, mode, and deletion
    \item \textbf{Outlier Detection}: Z-score based detection with configurable thresholds
    \item \textbf{Data Normalization}: Standardization of measurements for consistent analysis
\end{itemize}

\subsection{Error Handling and Validation}

Our implementation includes robust error handling and validation mechanisms:
\begin{itemize}
    \item \textbf{Input Validation}: Comprehensive checking of data structure and content
    \item \textbf{Type Safety}: Strict type checking and conversion
    \item \textbf{Error Logging}: Detailed logging with context and traceback information
    \item \textbf{Custom Exceptions}: Domain-specific error types for better error handling
\end{itemize}

\subsection{Implementation Details}

Our implementation focuses on:
\begin{itemize}
    \item \textbf{Modularity}: Each component is self-contained and follows consistent interfaces
    \item \textbf{Type Safety}: Comprehensive type hints and validation
    \item \textbf{Documentation}: Detailed docstrings and usage examples
    \item \textbf{Performance}: Optimized data structures and algorithms
    \item \textbf{Extensibility}: Easy addition of new scoring systems and data types
    \item \textbf{Testing}: Comprehensive test coverage for all components
\end{itemize}

\section{Results}

\subsection{Testing and Validation Results}

Our testing framework has demonstrated the reliability and robustness of the implementation:

\begin{itemize}
    \item \textbf{Synthetic Data Generation}: Successfully generated realistic MIMIC-like datasets with appropriate distributions and relationships
    \item \textbf{Component Validation}: All major components passed their respective validation tests
    \item \textbf{Integration Testing}: The complete pipeline successfully processed test data while maintaining data integrity
    \item \textbf{Performance Metrics}:
    \begin{itemize}
        \item Efficient handling of large datasets (10,000+ patients)
        \item Optimized memory usage with controlled growth
        \item Scalable processing speed with parallel processing capabilities
    \end{itemize}
\end{itemize}

\section{Discussion}
% Discussion content goes here

\section{Conclusion}
% Conclusion content goes here

\bibliographystyle{unsrtnat}
\bibliography{references}

\end{document}